Ricci-Ollivier curvature of two random walks on rooted phylogenetic subtree-prune-regraft graph

### 2015-02-19 curvature.tex
Typo.
Little twiddles in intro.

The SPR graph is very important in phylogenetic inference, and yet still little is known about the rooted or unrooted versions of the SPR graph itself, or random walks thereupon.
\cite{Song2003-gf} developed a recursive procedure on a tree to find the degree of the corresponding vertex in the rSPR graph.
\cite{Ding2011-bj} showed that the diameter $\Delta_{\text{rSPR}}$ of the rSPR graph is $n - \Theta(\sqrt n)$, and for the unrooted case they show
$$ n - 2\ceil{\sqrt{n}} + 1
\le \Delta_{\text{uSPR}}(n)
\le n - 3 - \floorV{\frac{\sqrt{n - 2} - 1}{2}}.
$$

Recently, a new approach to quantifying properties of random walks has been pioneered by Ollivier and colleagues \cite{Ollivier2009-bw,Joulin2010-jg}.
They define a notion of curvature on a discrete or continuous case.
Formalizes idea of the extent to which random walking brings points together, and is generalization of the notion of Ricci curvature on manifolds.

@@

\begin{lemma}
	The maximum curvature of the uniform random walk between two adjacent trees with $n$ leaves is $\frac{6n-17}{3n^2-13n+14}$. \todo{lazy random walk is $\frac{6n-15}{3n^2 - 13n + 14}$}
\end{lemma}
\begin{proof}
Let $T$ and $S$ be two adjacent trees.
The maximum curvature occurs when their neighborhoods have maximum overlap, are of equal size, and all other tree pairs are at distance 1.
By Lemma~\ref{lem:shared_neighbors} the maximum overlap is $6n-17$.
The maximum curvature is thus $1 - \frac{\degree{T} - 6n-17}{\degree{T}} = \frac{6n - 17}{\degree{T}}$. This is maximized when $\degree{T}$ is as small as possible, that is $T,S$ are ladders and $\degree{T} = 3n^2 - 13n + 14$.
\end{proof}

>>>>

The SPR graph is very important in phylogenetic inference, and yet still little is known about the rooted or unrooted versions of the SPR graph itself, or random walks thereupon.
\cite{Song2003-gf} developed a recursive procedure on a tree to find the degree of the corresponding vertex in the rSPR graph, and corresponding bounds on degree.
\cite{Ding2011-bj} showed that the diameter $\Delta_{\text{rSPR}}$ of the rSPR graph is $n - \Theta(\sqrt n)$, and for the unrooted case they show
$$ n - 2\ceil{\sqrt{n}} + 1
\le \Delta_{\text{uSPR}}(n)
\le n - 3 - \floorV{\frac{\sqrt{n - 2} - 1}{2}}.
$$
Despite the centrality of the SPR graph to phylogenetic inference, we are not aware of any other work investigating its properties.

Recently, a new approach to quantifying properties of discrete and continuous spaces with respect to random walks has been pioneered by Ollivier and colleagues \cite{Ollivier2009-bw,Joulin2010-jg}.
They define a notion of curvature on a discrete or continuous case.
Formalizes idea of the extent to which random walking brings points together, and is generalization of the notion of Ricci curvature on manifolds.

@@

\begin{lemma}
	The maximum curvature of the uniform random walk between two adjacent trees with $n$ leaves is $\frac{6n-17}{3n^2-13n+14}$. \todo{lazy random walk is $\frac{6n-15}{3n^2 - 13n + 14}$}
\end{lemma}
\begin{proof}
Let $T$ and $S$ be two adjacent trees.
The maximum curvature occurs when their neighborhoods have maximum overlap, are of equal size, and all other tree pairs are at distance 1.
By Lemma~\ref{lem:shared_neighbors} the maximum overlap is $6n-17$.
The maximum curvature is thus $1 - \frac{\degree{T} - (6n-17)}{\degree{T}} = \frac{6n - 17}{\degree{T}}$. This is maximized when $\degree{T}$ is as small as possible, that is $T,S$ are ladders and $\degree{T} = 3n^2 - 13n + 14$.
\end{proof}
